---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

inspect the first few lines of the data set:

what variables are there? of what type are the variables? what does each variable mean? what do we expect their data range to be?
is there a variable of special interest or focus?

Since this is on housing sales, sale price is likely a variable of interest.

16 variables, mainly quantitative.
Parcel ID
#Used to answer required questions
#summary(ames["Parcel ID"])
character with ID.
Type: Character
Length: 6935 (No Range values)

Address
property address in Ames, IA.
#Used to answer required questions
#summary(ames["Parcel ID"])
Type: character
Length: 6935 (No Range values)

Style
factor variable detailing the type of housing.
#Used to Answer required questions
#summary(ames['Style"])
Type: character
1 Story Frame: 3732
2 Story Frame: 1456
1 1/2 Story Frame: 711
Split Level Frame: 215
Split Foyer Frame: 156
Other: 218
NA: 447

Occupancy
factor variable of type of housing.
Type:Character
#Used to answer required question
#summary(ames["Occupancy"])
Condominium: 711
Single-Family/Owner occupied: 4711
Townhouse: 745
Two-Family Conversion: 139
NA's: 447

Sale Date
date of sale.

Sale Price
sales price (in US dollar).

Multi Sale
logical value: was this sale part of a package?

YearBuilt
integer value: year in which the house was built.

Acres
acres of the lot.

TotalLivingArea (sf)
total living area in square feet.

Bedrooms
number of bedrooms.

FinishedBsmtArea (sf)
total area of the finished basement in square feet.

LotArea(sf)
total lot area in square feet.
Quantitative- Range is between zero and 523228.

AC
logical value: does the property have an AC?
Expected- Yes or no, range- binary

FirePlace
logical value: does the property have an fireplace?
Expected yes or not, range- it is binary

Neighborhood
factor variable - levels indicate neighborhood area in Ames.
Not continuous, contains the specific neighborhood codes for the house sold
range- These are descrete values even though they are numbers

3. start the exploration with the main variable:

what is the range of the variable? draw a histogram for a numeric variable or a bar chart, if the variable is categorical. what is the general pattern? is there anything odd?
follow-up on oddities: see 4


```{r}
library(classdata)
library(ggplot2)

ames
range(ames$`Sale Price`)

ggplot(ames,
       aes(x = `Sale Price`)) +
  geom_histogram()
```
Most houses are on the lower end of the range, with a couple outliers in the upper millions. The data is skewed right and unimodal. This makes sense since there is an expected bottom to housing prices but mansions and large estates are expected to break the general trend when compared to average suburban housing. Several houses also have a sale price of zero, which may be an entry error.

4. pick a variable that might be related to the main variable. what is the range of that variable? plot. describe the pattern.
what is the relationship to the main variable? plot a scatterplot, boxplot or facetted barcharts (dependening on the types of variables involved). Describe overall pattern, does this variable describe any oddities discovered in 3? Identify/follow-up on any oddities.


4. Cameron Kraklio- Square feet compared to sale price
```{r}
library(classdata)
library(ggplot2)

ames
range(ames$`TotalLivingArea (sf)`)

ggplot(ames,
       aes(x = `TotalLivingArea (sf)`, y = `Sale Price`)) +
  geom_point()
```
It is a linear positive relationship. There are several outliers with small living areas but high prices. This could be due to a high lot area (like a farm house). This helps to explain the outliers that we saw earlier as they both have high sale price.
