---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.


inspect the first few lines of the data set:

what variables are there? of what type are the variables? what does each variable mean? what do we expect their data range to be?
is there a variable of special interest or focus?

Since this is on housing sales, sale price is likely a variable of interest.

16 variables, mainly quantitative.
Parcel ID
character with ID.

Address
property address in Ames, IA.

Style
factor variable detailing the type of housing.

Occupancy
factor variable of type of housing.

Sale Date
date of sale.

Sale Price
sales price (in US dollar).

Multi Sale
logical value: was this sale part of a package?

YearBuilt
integer value: year in which the house was built.

Acres
acres of the lot.

TotalLivingArea (sf)
total living area in square feet.

Bedrooms
number of bedrooms.

FinishedBsmtArea (sf)
total area of the finished basement in square feet.

LotArea(sf)
total lot area in square feet.
Quantitative- Range is between zero and 523228.

AC
logical value: does the property have an AC?
Expected- Yes or no, range- binary

FirePlace
logical value: does the property have an fireplace?
Expected yes or not, range- it is binary

Neighborhood
factor variable - levels indicate neighborhood area in Ames.
Not continuous, contains the specific neighborhood codes for the house sold

start the exploration with the main variable:

what is the range of the variable? draw a histogram for a numeric variable or a bar chart, if the variable is categorical. what is the general pattern? is there anything odd?
follow-up on oddities: see 4
pick a variable that might be related to the main variable.

```{r}
library(classdata)

ames
range(ames$`Sale Price`)
```


what is the range of that variable? plot. describe the pattern.
what is the relationship to the main variable? plot a scatterplot, boxplot or facetted barcharts (dependening on the types of variables involved). Describe overall pattern, does this variable describe any oddities discovered in 3? Identify/follow-up on any oddities
